Importance sampling applied to the simulation of extreme mechanical efforts exerted by a turbulent flow on a bluff body

### 2019-04-19 draft_p2.tex
Update text in part IV.B.2 Application of GKTL

We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag}.
Consistently with section~\ref{sec:time_avg}, we set the averaging window as $T=10\tau_c$.

@@

% Commentaire de la figure IS_GKTL
Figure~\ref{fig:illustration_IS} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N}\sum_{j=1}^{N}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the <<unbiased>> \ac{pdf} $\set{P}_0$ describing the time-averaged drag process. This estimate was computed on the basis of a control simulation of the flow spanning $T_{tot} = 10^6 \tau_c$ (see section~\ref{sec:pdfs}).
Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.

@@

For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
For $k \gtrapprox k_{max}$ the majority of the clones are resampled, with the expection of only one or two.
Because the resampling is based on a very small subset of the ensemble of $N_c$ trajectories, most of the trajectories in the biased ensemble overlap.
This effect is visible in figure~\ref{fig:illustration_IS}, where the histogram corresponding to the largest value of $k$ is peaked around a single value.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
See~\cite{ref1,ref2,ref3} for further details.

For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.

% Malgré les limitations, c'est toujours mieux
% qu'un échantillonnage direct
Sampling larger fluctuations therefore requires a larger population size, and a larger computational cost.
Nevertheless, table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample trajectories corresponding to extreme fluctuations of the time-averaged drag that could not be sampled through a direct sampling with equal computational cost.
For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T_a} \geq a$.
Note that trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.} are counted as one.  
As a comparison, it also displays the average number of such fluctuations in a control timeseries of duration $C_{gktl}$, divided in $N_c = C_{gktl} / T_a$ events.
To estimate this number we assumed that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.
Even though such a Gaussian assumption does not hold for fluctuations $F_T \geq 4\sigma_T$,
it may still provide numbers relevant to the comparison with the algorithm.
\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
  \caption{PDF of average drag over $10\tau_c$}
  \label{fig:PDF_AVG}
\end{figure}

In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
In the latter case, each fluctuation typically lasts a correlation time $\tau_c$.
In this context, the \ac{gktl} algorithm is particularly suited to sample trajectories that display a succession of these fluctuations.
(\textbf{conclusion à développer.})
\begin{table}
  \centering
  \begin{tabular}{l|ccccc}
     & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
    \hline
    $k=0.02$ & 1594 & 799 & 155 & 22 & 0  \\ 
    $k=0.025$ 1019 & 834 & 521 & 198 & 27  \\ 
    $k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
    $\mathcal{N}_{direct}$ & 2601.258& 374.862& 24.268& 0.958& 0.000\\ 
  \end{tabular}
  \caption{Number of independent corresponding to fluctuations $F_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the biased ensemble, for the three experiments $k=0.02,0.025,0.03$. Parameters of the \ac{gktl} experiments are $N_c = 16384$, $T_a=10\tau_c$, $\tau = \tau_c /2$. Within the biased ensemble, trajectories are considered distinct if they overlap for less than $50\%$ of their duration.
    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F'_{T_a} \geq a$ on can expect from $N_c$ independent realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.
    This number is based on the fact that the \ac{pdf} of $F_{T_a}$ is Gaussian.}
\label{tbl:nbTraj_tbl}
\end{table}

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
  \caption{Importance sampling for the time-averaged drag acting on the obstacle embedded in the test flow. The drag is averaged over $T=10\tau_c$. 
  The \ac{gktl} algorithm is used with $N_c = 16384$ copies and $T_a = T= 10\tau_c$ for three increasing values of the bias $k$. Additionally, the cloning period is $\tau = \tau_c/2$, resulting in $M=20$ iterations of the evolution/selection procedure.}
\end{figure}

>>>>

We now discuss the application of the \ac{gktl} algorithm to the flow dynamics introduced in section~\ref{sec:test_flow}.
The objective is to sample trajectories displaying extreme fluctuations of the \textit{time-averaged} drag~\eqref{eq:def_time_averaged_drag} acting on the square.
Consistently with section~\ref{sec:time_avg}, we set the averaging window as $T=10\tau_c$.

@@

% Commentaire de la figure IS_GKTL
Figure~\ref{fig:illustration_IS} illustrates the importance sampling achieved by the \ac{gktl} algorithm for the time-averaged drag $F_T$ with $T = T_a = 10\tau_c$.
For each of the three experiments, it displays the estimate of the biased \ac{pdf} of the time-average drag
\begin{equation}
  \label{eq:estimate_biased_measure}
  \rho_k(F) = \mathbb{E}_{\set{P}_k}[\delta(F[\{\mathbf{x}\}_{0\leq t \leq T_a}]-F)]
  \approx \frac{1}{N}\sum_{j=1}^{N_c}\delta(F^{(j)}_{T}[\{\mathbf{x}\}_{0\leq t \leq T_a}] - F),
\end{equation}
computed over the set of $N_c=16384$ trajectories sampled by the algorithm.
Moreover, figure~\ref{fig:illustration_IS_drag} displays an estimate of the unbiased \ac{pdf} describing the time-averaged drag process.
This estimate was computed on the basis the control run, see section~\ref{sec:pdfs}.
Note that the unbiased \ac{pdf} could also be estimated using the \ac{gktl} algorithm with $k=0$.

@@

For a fixed number of trajectories $N_c$, there is however an upper bound $k_{max}$ on the parameter $k$ above which finite size effects harm the efficiency and accuracy of the selection procedure.
For $k \gtrapprox k_{max}$ the resampling is based on a very small subset of the ensemble of trajectories, and most of the trajectories in the biased ensemble overlap.
This effect is visible in figure~\ref{fig:illustration_IS}, where the \ac{pdf} estimate for $k=0.03$ is peaked.
Note that in practice,there is no such sharp upper bound $k_{max}$: finite size effects become gradually relevant as the bias $k$ is increased.
An important question regarding the application of the \ac{gktl} algorithm is therefore that of the scaling of $k_{max}$ as a function of the ensemble size $N$. 
See~\cite{ref1,ref2,ref3} for further details.
% how to assess the value of k_max ?
For the application discussed in this paper, the order of magnitude of $k_{max}$ was assessed empirically, by assessing the diversity of the trajectories in the biased ensemble.
This diversity can be monitored at each selection stage of the algorithm, by computing the proportion of trajectories that share the same ancestor trajectory.

% Malgré les limitations, c'est toujours mieux
% qu'un échantillonnage direct
Table~\ref{tbl:nbTraj_tbl} shows that the \ac{gktl} algorithm is able to sample fluctuations of the time-averaged drag that could not be sampled through a direct sampling with equal computational cost.
For each three \ac{gktl} experiments, table~\ref{tbl:nbTraj_tbl} displays the number of trajectories that correspond to a time-average $F_{T} \geq a$.
Note that we count as one trajectories that overlap over more than $50\%$ of their total duration~\footnote{Trajectories that overlap over an interval $[O;t]$, with $t\geq T_a / 2$.}.
As a comparison, table~\ref{tbl:nbTraj_tbl} indicates the average number of fluctuations one can expect from simulating $N_c$ independant trajectories of duration $T_a$ $\{f_d(t)\}_{0\leq t \leq T_a}$, this with the same computational cost $C_{gktl}$.
This number assumes that the values of the time-average drag $F_T$, with $T=10\tau_c$, are Gaussian distributed.
This hypothesis is motivated by figure~\ref{fig:PDF_AVG}.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{PDF_AVG/PDF_AVG}
  \caption{PDF of average drag over $10\tau_c$}
  \label{fig:PDF_AVG}
\end{figure}

In section~\ref{sec:time_avg} we observed that extremely large values of the time-averaged drag can either result from a succession of rather typical---however all positive---fluctuations of the instantaneous drag $f_d$, or from a small number of extremely large fluctuations of the instantaneous drag.
In the latter case, each fluctuation typically lasts a correlation time $\tau_c$.
In this context, the \ac{gktl} algorithm is particularly suited to sample trajectories that display a succession of these fluctuations.
(\textbf{conclusion à développer.})
\begin{table}
  \centering
  \begin{tabular}{l|ccccc}
     & $\sigma$ & $2\sigma$ & $3\sigma$ & $4\sigma$ & $5\sigma$ \\
    \hline
    $k=0.02$ & 1594 & 799 & 155 & 22 & 0  \\ 
    $k=0.025$ 1019 & 834 & 521 & 198 & 27  \\ 
    $k=0.03$ & 539 & 510 & 391 & 205 & 36  \\ 
    $\mathcal{N}_{direct}$ & 2601.258& 374.862& 24.268& 0.958& 0.000\\ 
  \end{tabular}
  \caption{Number of fluctuations $F_{T_a} \geq a$ with a=$\sigma,2\sigma...$ in the biased ensemble, for the three experiments $k=0.02,0.025,0.03$.
    Within the biased ensemble, trajectories are counted as one if they overlap over more than $50\%$ of their duration.
    In addition, $\mathcal{N}_{direct}$ is the average number of fluctuations $F_{T_a} \geq a$ on can expect from $N_c$ independent realisations of $F_{T_a}$, \textit{i.e.} with $k=0$.}
\label{tbl:nbTraj_tbl}
\end{table}

\begin{figure}
  \centering
  \includegraphics[width=0.7\linewidth]{IS_GKTL/IS_GKTL}
  \caption{Importance sampling for the time-averaged drag $F_T$ with $T=10\tau_c$. The shaded \ac{pdf} estimates are computed on the biased ensemble resutling from the \ac{gktl} algorithm.
The dashed line denotes the unbiased \ac{pdf} describing $F_T$, \textit{i.e.} statistics sampled without the \ac{gktl} algorithm, or with $k=0$ (no bias)}
\end{figure}
